\documentclass[a4paper]{book}
\input{preliminaries.tex}
\begin{document}


\begin{itemize}
\item ``Not a single Greek should be left in Istanbul until the 500th anniversary of its conquest"; in an official report called ``Minority Report", by CHP (\ie the founding party). (\#7)
\item Most of remaining churches in today's Turkey are built after 1839 (Tanzimat) (\#29)
\item There were 288 Greek schools in Trabzon by 1898! (p46)
\item English lieutenant L.H. Hurst reported as early as 1919 that ``Mustafa Kemal is preparing for activities that will cause Christian slaughter/pogrim (turkish word: kiyim) p52"
\item Topal Osman, the infamous gang leader responsible for Armenian and Greek massacres, who was wanted by Istanbul Divan-i Harbi for his crimes during Armenian Genocide, was pardoned and appointed into an official post with Mustafa Kemal's special order! p52
	\begin{itemize}
	\item The same Topal Osman proudly admitted that he murdered Greeks and other non-Muslims and appropriated their properties. p52 
	\item The Greeks who were deported in 1921 were killed in masses by Topal Osman and his gangs on their way to deportation. p52
	\item The same Osman was later invited by Mustafa Kemal to the special Bodyguard Regimen.
	\item The same Osman later murdered the Trabzon Deputy. Mustafa Kemal ordered Osman to be taken into custody; he was captured and then executed. A day after his execution, his corpse was dug out and decapitated and publicly displayed.
	\item Topal Osman's statue was erected in Turkey in ... 2007.
	\end{itemize}
\item The Aegean cost was ``Turkified" before the war by forced deportation. Numbers vary between 150,000 and 1,150,100 p125-127.
\end{itemize}

\begin{itemize}
\item Apparently the origins of Karamanlides is highly disputed 
\end{itemize}

\begin{itemize}
\item To break the influence of Fener Rum Patriarchy, a Turkish (!) Patriarchy was founded by a ``Papa Eftim" (later Zeki Erenerol !) p64.
\item Ottoman Empire forbade the printing of the bible in 1883. (Technically, they first asked a Lewis Wallace to add to the front page the note ``Only for the usage of Protestants'', which he refused to do saying that this was something that was never done in the history of the bible, p74.
\item \textbf{Assyrians.} The story of Assyrians is particularly interesting. They consistently declared their support to the Ottomans and the Turks (p.113 ), yet weren't favored by the latter (p119-120).
\end{itemize}




\bibliographystyle{IEEEtran}
\bibliography{../bibliography}

\end{document}